Dependent Object Types

### 2012-08-10 article.tex
compromise

\subsubsection{Set of Values}
$\relv k \Gamma s T$ defines the set of values that appear to have
type $T$ when taking at most $k$ steps. $\Gamma$ and $s$ must agree:
$\dom(\Gamma) = \dom(s)$ (ordered) and $\forall (x : T) \in \Gamma, x
\in \relv k \Gamma s T$. A variable $y$ belongs to $\relv 0 \Gamma s
T$ simply by being in the store. In addition, it belongs to $\relv k
\Gamma s T$ for $k > 0$, if it defines all type, method and value labels in
the expansion of $T$ appropriately for $j < k$ steps.
\begin{align*}
&\relv k \Gamma s T = \{ y \ |\\
&\gap \Gamma \ts T \wfe \andl\\
&\gap y \in \dom(s) \andl\\
&\gap \forall j < k,\\
&\gap y \mapsto T_c \ldefs{\seq{l = v}\;\seq{m(x)=t}} \in s,\\
&\gap \Gamma \ts T \expand_y \seq{D},\\
&\gap (\forall L_i : S \rightarrow U \in \seq{D},\\
&\gap\gap \Gamma \ts y.L_i \wfe \spcomma S \wf \spcomma U \wf \spcomma S \sub U \andl\\
&\gap\gap \Gamma \ts y \ni L_i \typ {S'..U'} \spcomma S \sub S' \spcomma U' \sub U\\
&\gap (\forall m_i : S \rightarrow U \in \seq{D},\\
&\gap\gap \Gamma \ts S \wf \spcomma U \wf \andl t_i \in \rele j {\Gamma \envplus{x_i : S}} s U)\\
&\gap \andl\\
&\gap (\forall l_i : V \in \seq{D},\\
&\gap\gap \Gamma \ts V \wf \andl v_i \in \relv j \Gamma s V)\\
&\}
\end{align*}

This relation captures the observation that the only ways for a term
to get stuck is to have a field selection on an uninitialized field or
a method invocation on an uninitialized method. However, a {\it potential
  pitfall} is that the value itself might occur in the types $S$, $U$,
$V$, because we substitute it for the ``self'' occurrences in the
expansion, so the relation makes sure that the type labels are appropriate.

@@

\begin{proofsketch}
By (1) and (3), it should hold that (5) $v \in \relv {k_2}
{\Gamma_{12}} {s_{12}} {S}$ by the subset semantics lemma.  Since (2)
holds, it should also hold that ${t} \in \rele {\min(k_1,k_2)}
{\Gamma_{12} \envplus{{x} : S}} {s_{12}} T$ by the extended
monotonicity lemma. Then, we can instantiate $x$ in the complete store
to map to what $v$ maps to. This should be fine by (5) and
monotonicity. Thus, ${t} \in \rele {\min(k_1,k_2)} {\Gamma_{12}
  \envplus{{x} : S}} {s_{12} \envplus{{x} \mapsto s_{12}(v)}}
T$. Thanks to (4), we don't actually need $x$ to be held abstract in
the environment, because it won't occur in $T$ or its expansion, so we
can use the type of $v$ in the environment instead of
$S$ for $x$: ${t} \in \rele {\min(k_1,k_2)} {\Gamma_{12} \envplus{{x}
    : {\Gamma_{12}(v)}}} {s_{12} \envplus{{x} \mapsto s_{12}(v)}} T$. This
implies what needs to be shown.

\end{proofsketch}

@@

A {\it potential pitfall} is whether some types of the expansion of
$U$ can become non-expanding when the ``self'' occurrences are
actually $v$ instead of just abstractly of type $S$, causing a check
to fail. Another worry is that such a non-expanding type results from
narrowing of a parameter type. However, we believe this is fine by
subtyping regularity wtr to $\nswfe$ and $v \in \relv k \Gamma s S$
which cheks bounds and $\nswfe$ of the type labels.
\end{proofsketch}

>>>>

\subsubsection{Set of Values}
$\relv k \Gamma s T$ defines the set of values that appear to have
type $T$ when taking at most $k$ steps. $\Gamma$ and $s$ must agree:
$\dom(\Gamma) = \dom(s)$ (ordered) and $\forall (x : T) \in \Gamma, x
\in \relv k \Gamma s T$. A variable $y$ belongs to $\relv 0 \Gamma s
T$ simply by being in the store. In addition, it belongs to $\relv k
\Gamma s T$ for $k > 0$, if it defines all type, method and value labels in
the expansion of $T$ appropriately for $j < k$ steps.
\begin{align*}
&\relv k \Gamma s T = \{ y \ |\\
&\gap \Gamma \ts T \wfe \andl\\
&\gap y \in \dom(s) \andl\\
&\gap \forall j < k,\\
&\gap y \mapsto T_c \ldefs{\seq{l = v}\;\seq{m(x)=t}} \in s,\\
&\gap \Gamma \ts T \expand_y \seq{D},\\
&\gap (\forall L_i : S \rightarrow U \in \seq{D},\\
&\gap\gap \Gamma \ts y \ni L_i \typ {S'..U'}) \andl\\
&\gap (\forall m_i : S \rightarrow U \in \seq{D},\\
&\gap\gap t_i \in \rele j {\Gamma \envplus{x_i : S}} s U) \andl\\
&\gap (\forall l_i : V \in \seq{D},\\
&\gap\gap v_i \in \relv j \Gamma s V)\\
&\}
\end{align*}

This relation captures the observation that the only ways for a term
to get stuck is to have a field selection on an uninitialized field or
a method invocation on an uninitialized method. However, a {\it potential
  pitfall} is that the value itself might occur in the types $S$, $U$,
$V$, because we substitute it for the ``self'' occurrences in the
expansion, so the relation makes sure that the type labels exist.

@@

\begin{proofsketch}
By (1) and (3), it should hold that (5) $v \in \relv {k_2}
{\Gamma_{12}} {s_{12}} {S}$ by the subset semantics lemma.  Since (2)
holds, it should also hold that ${t} \in \rele {\min(k_1,k_2)}
{\Gamma_{12} \envplus{{x} : S}} {s_{12}} T$ by the extended
monotonicity lemma. Then, we can instantiate $x$ in the complete store
to map to what $v$ maps to. This should be fine by (5) and
monotonicity. Thus, ${t} \in \rele {\min(k_1,k_2)} {\Gamma_{12}
  \envplus{{x} : S}} {s_{12} \envplus{{x} \mapsto s_{12}(v)}}
T$. Thanks to (4), we don't actually need $x$ to be held abstract in
the environment, because it won't occur in $T$ or its expansion (a
{\it potential pitfall} is whether its occurrences in $t_i$ could
still cause a check to fail through narrowing issues), so we can use
the type of $v$ in the environment instead of $S$ for $x$: ${t} \in
\rele {\min(k_1,k_2)} {\Gamma_{12} \envplus{{x} : {\Gamma_{12}(v)}}}
      {s_{12} \envplus{{x} \mapsto s_{12}(v)}} T$. This implies what
      needs to be shown.
\end{proofsketch}

@@

A {\it potential pitfall} is whether some types of the expansion of
$U$ can become non-expanding when the ``self'' occurrences are
actually $v$ instead of just abstractly of type $S$, causing a check
to fail. Another worry is that such a non-expanding type results from
narrowing of a parameter type.
%However, we believe this is fine by
%subtyping regularity wtr to $\nswfe$ and $v \in \relv k \Gamma s S$
%which cheks bounds and $\nswfe$ of the type labels.
\end{proofsketch}
